Rapid response of mesophyll conductance to light availability allows shade leaves to take advantage of sunflecks

### 2015-11-30 manuscript/manuscript_wtc3.Rmd
comments from MT

# Summary
* Light gradients within tree canopies play a major role in the distribution of plant resources that define photosynthetic capacity of individual leaves. The transient nature of light availability within a canopy, however, complicates our ability to assess the contribution of shade leaves to canopy carbon gain.
* Leaf gas exchange was coupled with online carbon isotope discrimination to measure photosynthesis (A), stomatal conductance (g~s~) and mesophyll conductance (g~m~) of sun and shade leaves in *Eucalyptus tereticornis* trees grown in climate controlled whole tree chambers.
* Compared to sun leaves, shade leaves had lower A, g~m~, Vc~max~, J~max~ and leaf nitrogen but maintained similar rates of g~s~. When light intensity was increased from low light to high light for shade leaves both g~s~ and g~m~ increased rapidly, leading to increases in A greater than sun leaves at the same PPFD. 
* Here we show that dynamic physiological responses of shade leaves to altered light environments has important implications for upscaling leaf level measurements and predicting whole canopy carbon gain. We argue that the rapid response of g~m~ with light enables shade leaves to respond quickly to sunflecks. Evidence that g~m~ not only varies within a canopy but can be up-regulated over short time intervals possibly represents a new mechanism underpinning leaf gas exchange responses to light.

@@

\
Leaf level gas exchange was measured with a standard (2 x 3 cm) leaf cuvette using a portable gas exchange system (LI-6400XT, LI-COR, Lincoln, NE, USA).  This system was coupled with a tunable diode laser (TDL; TGA100,Campbell Scientific, Inc., Logan, UT, USA) for concurrent measurements of online C isotope discrimination. The CO~2~ in the leaf cuvette was set at ambient atmospheric [CO~2~] (400 ppm) with a flow rate of 200 $\mu$mol s^-1^. Two identical gas exchanges systems were run simultaneously, one in each of a randomly chosen WTC for each temperature treatment. Leaf temperatures were controlled at the current AT or ET WTC air temperature. PPFD in the cuvette was set to match the individual light environment of each leaf type (explained above). Sun flecks were simulated for shade leaves by increasing the leaf cuvette PPFD (LI-COR red/blue light source) to match the light environment of the full sun leaf in the same tree. The maximum sunfleck response was recorded once CO~2~ and water vapor fluxes re-stabilized in the leaf cuvette (ca. 25 min).

@@

\
Mean g~m~ was higher in sun compared to shade leaves (+27%) under their local light environment (P < 0.001). Following a short-term increase in PPFD from low to high light, g~m~ of shade leaves increased to values significantly greater than sun leaves (Table 2). Proportional increases in g~m~ were matched by proportional increases in A from low to high light in shade leaves (Figure `r gebarfig`b,c). Photosynthesis scaled positively with increases in g~m~ for all leaves and the large increases in g~m~ in shade leaves under high light resulted in the highest rates of A (Figure `r agmgsfig`b). No differences in g~m~ were detected with the warming treatment within leaf types. Mesophyll conductance did not vary across measurements campaigns within leaf types and light treatments, but a weak negative relationship with increasing leaf temperature was detected with leaf types under local light conditions (Figure `r get1fig`b). We also simulated AC~c~ curves to determine if treatment differences in J~max~ and Vc~max~ where instead the result of differences in g~m~. Comparison of A-Cc curves (Figure `r accfig`) and A-Ci curves revealed similar differences between sun and shade leaves.

@@

# Discussion
Here we show that leaves within canopies of *Eucalyptus tereticornis* trees maintain the ability to upregulate A beyond the limitations of their local light environment.

More open stomata and rapid increases in g~m~ allowed shade leaves to readily utilize increases in light availability, leading to rates of leaf C gain exceeding sun leaves.

Although we know shade leaves experience transient periods of sun and shade [@pearcy1990sunflecks], these findings are unexpected because a lack of empirical data within tree canopies currently impedes our ability to predict whole canopy C gain.

This dynamic behaviour of shade leaves complicates optimizations of canopy C uptake and accounting for the response of leaf physiology to transient light dynamics is an essential need in future studies.

Additionally, with measurements recorded across a large natural range of air temperatures only minimal effects a +3 &deg;C warming treatment were detected on leaf physiology.

@@

\
If shade leaves "lie in wait" for sunflecks then perhaps we should consider an alternate leaf economic strategy to maximize C gain, beyond conventional trade-offs associated with canopy resource distribution. If the valuation of sunflecks as a C resource is large enough then costs of sub-optimal stomatal behaviour could be offset over the leaf lifespan. For example, the potential C gain in leaves where sunflecks constitute a large proportion of total daily PFFD may be large enough to tolerate decreases in WUE. However, accounting for the heterogeneous nature of light within a canopy remains a current challenge for empirical and modelling studies. Thus, models which predict leaf photosynthesis from N distribution within a canopy will be incomplete unless inclusion of canopy light extinction and the integration of sunflecks on shade leaves are included [@de1997simple].

>>>>

# Summary
* Light gradients within tree canopies play a major role in the distribution of plant resources that define photosynthetic capacity of individual leaves. A lack of empirical data relating photosynthesis to leaf physiolgoical behavior within tree canopies, however, impedes our ability to assess the contribution of shade leaves to canopy carbon gain.
* To investigate the CO~2~ diffusion pathway of sun and shade leaves, leaf gas exchange was coupled with online carbon isotope discrimination to measure photosynthesis (A), stomatal conductance (g~s~) and mesophyll conductance (g~m~) in *Eucalyptus tereticornis* trees grown in climate controlled whole tree chambers.
* Compared to sun leaves, shade leaves had lower A, g~m~, Vc~max~, J~max~ and leaf nitrogen but maintained similar rates of g~s~. When light intensity was increased from low light to high light for shade leaves both g~s~ and g~m~ increased rapidly, leading to increases in A greater than sun leaves at the same PPFD. 
* Here we show that dynamic physiological responses of shade leaves to altered light environments has important implications for upscaling leaf level measurements and predicting whole canopy carbon gain. We argue that the rapid response of g~m~ with light enables shade leaves to respond quickly to sunflecks. Evidence that g~m~ not only varies within a canopy but can be up-regulated over short time intervals possibly represents a new mechanism underpinning leaf gas exchange responses to light.

@@

\
Leaf level gas exchange was measured with a standard (2 x 3 cm) leaf cuvette using a portable gas exchange system (LI-6400XT, LI-COR, Lincoln, NE, USA).  This system was coupled with a tunable diode laser (TDL; TGA100,Campbell Scientific, Inc., Logan, UT, USA) for concurrent measurements of online C isotope discrimination. The CO~2~ in the leaf cuvette was set at ambient atmospheric [CO~2~] (400 ppm) with a flow rate of 200 $\mu$mol s^-1^. Two identical gas exchanges systems were run simultaneously, one in each of a randomly chosen WTC for each temperature treatment. Leaf temperatures were controlled at the current AT or ET WTC air temperature. PPFD in the cuvette was set to match the individual light environment of each leaf type (explained above). Periods of high irradiance were simulated for shade leaves by increasing the leaf cuvette PPFD (LI-COR red/blue light source) to match the light environment of the full sun leaf in the same tree. The maximum sunfleck response of shade leaves was then recorded once CO~2~ and water vapor fluxes re-stabilized in the leaf cuvette (ca. 25 min).

@@

\
Mean g~m~ was higher in sun compared to shade leaves (+27%) under their local light environment (P < 0.001). Following a short-term increase in PPFD from low to high light, g~m~ of shade leaves increased to values significantly greater than sun leaves (Table 2). Proportional increases in g~m~ were matched by proportional increases in A from low to high light in shade leaves (Figure `r gebarfig`b,c). Photosynthesis scaled positively with increases in g~m~ for all leaves, with similar intercepts but different slopes between leaf type and light treatment (P = 0.0186). The large increases in g~m~ in shade leaves under high light resulted in the highest rates of A (Figure `r agmgsfig`b). No differences in g~m~ were detected with the warming treatment within leaf types. Mesophyll conductance did not vary across measurements campaigns within leaf types and light treatments (Figure `r get1fig`b), but a weak negative relationship with increasing leaf temperature was detected with sun and shade leaves under their local light environment (P = 0.001 & 0.04, respectively). We also simulated AC~c~ curves to determine if treatment differences in J~max~ and Vc~max~ where instead the result of differences in g~m~. Comparison of A-Cc curves (Figure `r accfig`) and A-Ci curves revealed similar differences between sun and shade leaves.

@@

# Discussion
Here we show that A in leaves within canopies of *Eucalyptus tereticornis* are limited by their local light environment, however, shade leaves increased rates of leaf C gain exceeding sun leaves when light availability increased. Although shade leaves in lower light environments exhibited relatively high g~s~ and c~i~, it was rapid increases in g~m~ under periods of high light availability that led to this upregulation of A. Although we know shade leaves experience transient periods of sun and shade [@pearcy1990sunflecks], a lack of empirical data within tree canopies currently impedes our ability to predict whole canopy C gain. These findings offer new insights into how aspects of leaf physiology may be optimized differently in sun and shade leaves and reveal how the total leaf CO~2~ conductance pathway should be accounted for when testing optimizations of canopy C uptake in future studies. Additionally, with measurements recorded across a large natural range of air temperatures only minimal effects a +3 &deg;C warming treatment were detected on leaf physiology.

@@

\
If shade leaves "lie in wait" for sunflecks then perhaps we should consider an alternate leaf economic strategy to maximize C gain, beyond conventional trade-offs associated with canopy resource distribution. This is because the role of g~s~ in regulating photosynthetic induction impacts the capacity of a leaf to utilize sunflecks [@way2012sunflecks]. If the valuation of sunflecks as a C resource is large enough then costs of sub-optimal stomatal behaviour could be offset over the leaf lifespan. For example, the potential C gain in leaves where sunflecks constitute a large proportion of total daily PFFD may be large enough to tolerate decreases in WUE. However, accounting for the heterogeneous nature of light within a canopy remains a current challenge for empirical and modelling studies. Thus, models which predict leaf photosynthesis from N distribution within a canopy will be incomplete unless inclusion of canopy light extinction and the integration of sunflecks on shade leaves are included [@de1997simple].
